Predictions for detecting BHNS and other double compact objects binaries with LISA

### 2021-04-18 paper/tex/sections/introduction.tex
update BHBH section of intro

Double black hole (BHBH) binaries are the dominant source in ground-based gravitational wave detections and several studies have tried to model the black hole mass function as well as identify the pair instability mass gap \citep[e.g.][]{Baxter+2021}. In contrast, BHBHs in LISA are not biased towards higher masses and instead will probe lower mass black holes, potentially offering insight about the presence of a lower mass gap between neutron stars and black holes.

@@

One particularly interesting and elusive gravitational wave source is a black hole neutron star binary (BHNS). Of all the events detected by ground-based detectors, none can be confidently attributed to the merger of a black hole and a neutron star, though several events such as GW190425 and GW190814 have not been ruled out as a BHNS merger \citep{Abbott+2020_GW190425,Abbott+2020_GW190814}. Predictions for the merger rate of BHNSs range across three orders of magnitude \citep[e.g.][]{Broekgaarden+2021} so the number of detections in LISA will be important in reducing this uncertainty, thereby refining our understanding of the remnants and evolution of massive stars. These binaries are expected to have similar electromagnetic counterparts to NSNSs and so can studied in the same way. A distinctly exciting possibility is the detection of a pulsar--BH system or millisecond pulsar--BH system as these could be observed not only by gravitational wave detectors, but also radio telescopes such as MeerKAT and SKA, which will help to constrain uncertain binary evolution processes \citep{Chattopadhyay+2020}.

@@

We build upon previous efforts but will several important improvements. We explore the effect of varying binary physics assumptions by repeating our analysis for 15 different models and comparing the effect on the detection rate and distributions of source parameters. We use a model for the Milky Way that is dependent on the chemical enrichment history and calibrated on the latest GAIA and APOGEE surveys \citep{Frankel+2018}. In contrast to many previous works, we provide a full treatment of the eccentricity of detectable sources both for the inspiral evolution as well as gravitational wave signal during the LISA mission. Moreover, our binary population synthesis simulation is the most extensive of its kind, with 750 million binaries (one million binaries for each of 50 metallicity bins and 15 physics variations) evolved to produce the DCO populations used in this work \citep{Broekgaarden+2021}. In addition, we use the adaptive sampling algorithm STROOPWAFEL to further reduce our sampling noise \citep{Broekgaarden+2019}.

>>>>

For the purposes of this paper, we focus only on the classical isolated binary evolution channel \citep[e.g.][]{Tutukov+1973,Tutukov+1993,Kalogera+2007,Belczynski+2016} in which compact objects are formed through through highly non-conservative mass transfer or common envelope ejection. There are however several alternative proposed formation channels including: dynamical formation in dense star clusters \citep[e.g.][]{Sigurdsson+1993,PortegiesZwart+2000,Miller+2009,Rodriguez+2015} and (active) galactic nuclei discs \citep[e.g.][]{Morris+1993, Antonini+2016, McKernan+2020}, isolated hierarchical triple evolution involving Kozai-Lidov oscillations \citep[e.g.][]{Stephan+2016, Silsbee+2017,Antonini+2017} and chemically homogenous evolution through efficient rotational mixing \citep[e.g.][]{deMink+2009, deMink+2016,Marchant+2016,duBuisson+2020}.

@@

Double black hole (BHBH) binaries in the Milky Way provide a greater observational challenge and, at the time of writing, no system of this nature has been identified. Thus, LISA could provide the first detection of a Galactic BHBH binary. Moreover, the mass distribution of stellar mass black holes is still uncertain. The only confirmed black holes (BH) in our galaxy have been discovered as components of X-ray binaries with companion stars \citep[e.g.][]{Bolton+1972,Webster+1972}. This sample of BHs has masses mainly constrained between $5$ and $10 \unit{M_\odot}$ \citep{Corral-Santana+2016}, a stark contrast to the more massive BHs observed with LIGO/Virgo that tend to have masses concentrated around $30 \unit{M_{\odot}}$ \citep{Abbott+2020_GWTC2}. The mass distribution of Galactic BHs has been investigated in detail, however the presence of a lower mass gap (from $2$-$5 \unit{M_{\odot}}$) remains an open question \citep[e.g.][]{Ozel+2010,Farr+2011}. More recently there has been increased discussion over the maximum BH mass in our galaxy, with the claims of a $70 \unit{M_{\odot}}$ BH \citep{Liu+2019,Abdul-Masih+2020} and revised measurements of the mass of Cygnus X-1 \citep{Miller-Jones+2021}. A sample of BHBHs detected with LISA could possibly help to constrain these uncertainties in the stellar mass BH mass distribution.

One particularly interesting and elusive gravitational wave source is a black hole neutron star binary (BHNS). Of all the events detected by ground-based detectors, none can be confidently attributed to the merger of a black hole and a neutron star, though several events such as GW190425 and GW190814 have not been ruled out as a BHNS merger \citep{Abbott+2020_GW190425,Abbott+2020_GW190814}. Predictions for the merger rate of BHNSs range across three orders of magnitude \citep[e.g.][]{Abadie+2010, Broekgaarden+2021} so the number of detections in LISA will be important in reducing this uncertainty, thereby refining our understanding of the remnants and evolution of massive stars. These binaries are expected to have electromagnetic counterparts that can studied in the same way as NSNSs. A distinctly exciting possibility is the detection of a pulsar--BH system or millisecond pulsar--BH system \citep{Narayan+1991}. These systems could be observed not only by gravitational wave detectors, but also radio telescopes such as MeerKAT and SKA, which will help to constrain uncertain binary evolution processes \citep[e.g.][]{Pfahl+2005,Chattopadhyay+2020}.

@@

We build upon previous efforts but with several important improvements. We explore the effect of varying binary physics assumptions by repeating our analysis for 15 different models and comparing the effect on the detection rate and distributions of source parameters. We use a model for the Milky Way that is dependent on the chemical enrichment history and calibrated on the latest GAIA and APOGEE surveys \citep{Frankel+2018}. In contrast to many previous works, we provide a full treatment of the eccentricity of detectable sources both for the inspiral evolution as well as gravitational wave signal during the LISA mission. Moreover, our binary population synthesis simulation is the most extensive of its kind, with 750 million binaries (one million binaries for each of 50 metallicity bins and 15 physics variations) evolved to produce the DCO populations used in this work \citep{Broekgaarden+2021}. In addition, we use the adaptive sampling algorithm STROOPWAFEL to further reduce our sampling noise \citep{Broekgaarden+2019}.
